---
title: "Exploratory Data Analysis"
format: pdf
editor: visual
---

### Data

```{r, echo = FALSE, message = FALSE, warning = FALSE}
#Libraries and the Dataset
pacman::p_load(fpp3, patchwork)

read.csv(
  here::here("AnnualFertilityData_Cleaned.csv"), check.names = FALSE
) |>
  tsibble(index = Year) |>
  mutate(TLB_s = TLB / 1e4) -> fertility_df
```

The dataset has been provided by the Singapore Department of Statistics. It primarily covers births and fertility rates within Singapore, collecting data concerning a variety of variables from 1960 to 2025. However for the purpose of this data analysis the main variables that will be considered are:

-   `Year`:

    -   The EDA will be concerned with the entire period of 1960 to 2025. However for modelling and forecasting, it pertains to the periods 1960 to 2012, and 2013 to 2024 respectively.

-   `TLF`: Total Fertility Rate

    -   This refers to the average number of live-births each female would have during her reproductive years, subjected to prevailing age-specific fertility rates in the population in the given year.

    -   Additionally data prior to 1980 pertains to total population while onwards pertains to resident population (Singaporean citizens and permanent residents.

-   `TLB`: Total Live Births

    -   Data is concerned with all live births within Singapore and its territorial waters as registered by the Singaporean government at the year of birth.

-   `TLB_s`: Total Live Births scaled

    -   This is an additional variable introduced. It is a down scaling of the live births data that will be used for ease of analysis.

    -   E.g. If TLB were recorded 10,000 at the time then it's down scale would be 1.

```{r, echo = FALSE, message = FALSE}
head(fertility_df) |>
  select(Year, TFR, TLB, TLB_s) |>
  knitr::kable(caption = "Table 1: The first 6 rows of the dataset for display")
```

#### Summary Statistics

-   Total Fertility rate on average is 1.52 with with a standard deviation of 1.23. This is relatively high compared to the mean. So fertility rate increases increases and decreases per year by this amount.

```{r, echo = FALSE, message = FALSE}
fertility_df |>
  features(TFR, list(Mean = mean, 
                     Variance = var,
                     "Standard Deviation" = sd,
                     quantile)) |>
  mutate_if(is.numeric, ~ as.character(signif(.,3))) |>
  knitr::kable(caption = "Table 2: Summary statistics of the Total Fertility Rate")
```

-   On average the total live births per year is 4.41 or 44,100 with a standard deviation of 0.681. This standard deviation is lower than the mean. Total live births each year are increased and decreased by this amount or by 6,810.

```{r, echo = FALSE, message = FALSE}
fertility_df |>
  mutate(TLB_s = TLB / 1e4) |>
  features(TLB_s, list(Mean = mean,
                     Variance = var,
                     "Standard Deviation" = sd,
                     quantile)) |>
  mutate_if(is.numeric, ~ as.character(signif(.,3))) |>
  knitr::kable(caption = "Table 4: Sumamry statistics of the scaled Total Live Births")
```

#### Box-plots

```{r, echo = FALSE, message = FALSE}
fertility_df |>
  ggplot(aes(x = TFR, fill = after_stat(width))) +
  geom_boxplot() +
  ggsci::scale_fill_material("blue") +
  theme_light() +
  theme(legend.position = "none") -> b1

fertility_df |>
  ggplot(aes(x = TLB_s, fill = after_stat(width))) +
  geom_boxplot() +
  ggsci::scale_fill_material("blue") +
  theme_light() +
  theme(legend.position = "none") -> b2

b1 / b2
```

The box-plot Total Fertility Rate has no skewed with 10 outliers. Suggesting overall lower fertility rate is recorded. 50% of recorded fertility rates lies roughly between 1.5 and 2. The lowest fertility rate is 0.87. Maximum fertility rate is 5.76 as all seen in Table 1. This is maximum is an outlier and supported by the box-plot. In addition about half the recorded fertility rates lie below or above the median; 1.62. This is also seen on fertility rate box-plot.

Total Live Births box-plot is right skewed with no outliers. 50% of recorded total live births lie roughly between 4 and the low 5s. The recorded median is 4.24 seen in Table 2. Half lie below and above this median. The maximums and minimums here aren't outliers but are recorded as 6.18 and 3 respectively in Table 2. Essentially at a one point in time, the maximum total live births was 61,800 and the minimum was 30,000.

#### Histograms

Histogram of the Total Fertility Rate is unimodal right-skewed. A majority of recorded data leans towards the lower end of fertility rate. As reflected by the box-plot there are also clusters of outliers here beyond the TFR of 3.

```{r, echo = FALSE, message = FALSE}
fertility_df |>
  ggplot(aes(x = TFR, fill = after_stat(count))) +
  geom_histogram() + 
  ggsci::scale_fill_material("blue") +
  geom_histogram(col = "black") +
  theme_light() +
  theme(legend.position = "none")
```

The histogram ofTotal Live Births displays a bimodal distribution. There's two main peaks are 4 and 5. Additionally there no data in the centre between 4 and 5 separating the two peaks.

```{r, echo = FALSE, message = FALSE}
fertility_df |>
  ggplot(aes(x = TLB_s, fill = after_stat(count))) +
  geom_histogram() +
  ggsci::scale_fill_material("blue") +
  geom_histogram(col = "black") +
  theme_light() +
  theme(legend.position = "none")
```

#### Time Plots and Correlograms

```{r, echo = FALSE, message = FALSE}
fertility_df |>
  gg_tsdisplay(TFR, plot_type = 'partial')
```

The time series for total fertility rate, it's time plot has a decreasing trend. As it's annual data there is no seasonality. It's correlogram has a gradual decay. Past lag 10 the ACF is no longer significant. The partial correlogram cuts off sharply after lag 1. No more lags after have a significant PACF. Overall this suggests an autoregressive model could be considered. However, because of it's non-stationarity and decreasing trend this isn't a definitive conclusion. Since there's a possibility of the trend influencing these correlogram. Transformation and differencing should be considered to ensure stationarity.

```{r, echo = FALSE, message = FALSE}
fertility_df |>
  model(stl = STL(TFR)) |>
  components() |>
  autoplot()
```

The time decomposition of the total fertility rate further supports that it's a decreasing trend with no stationarity. The fluctuations seen in the remainder explains for the peaks and troughs seen in the time series. What these fluctuations could be is a potential avenue for further exploration. Additionally in the trend, it's clearly seen that it could be of an exponential decrease. So a logarithmic transform could be a consideration.

```{r, echo = FALSE, message = FALSE}
fertility_df |>
  gg_tsdisplay(TLB_s, plot_type = 'partial')
```

Total Live Births, it's time plot displays a similar non-seasonal decreasing trend. In addition, both time plots display a peak increase in their respective variables, for the periods between the 1980s and 2000s. This could be an indication of socio-economical features not present in this data that could worth investigating. The correlogram features an ACF that decays gradually and is not significant after lag 6. It's partial correlogram, a significant peak in it's PACF and cuts off afterwards. Suggesting a similar consideration of autoregressive and a similar conclusion that the trend is influencing these.

```{r, echo = FALSE, message = FALSE}
fertility_df |>
  model(stl = STL(TLB_s)) |>
  components() |>
  autoplot()
```

The Total Live Births decomposition shows that the fluctuations seen in the time plot are explained by a remainder, leaving only a decreasing trend. This noticeable supports the the conclusion that between 1980 and 2000 there was an increase in live births bucking the trend.

Overall there's a correlation between fertility rate and live births. It's relationship could be worth considering if one is an significant indicator of the other.

#### Modelling

```{r, echo = FALSE, message = FALSE}
fertility_df |>
  filter(Year <= 2012) -> train_df
```

```{r, echo = FALSE, message = FALSE}
fertility_df |>
  autoplot(log(TFR)) -> p1

fertility_df |>
  autoplot(TFR) -> p2

p1 / p2
```

```{r, echo = FALSE, message = FALSE}
train_df |>
  features(TFR, list(
    unitroot_kpss,
    unitroot_ndiffs)) |>
  mutate_if(is.numeric, ~ as.character(signif(.,3))) |>
  knitr::kable(caption = "Table 5: ")
```

```{r, echo = FALSE, message = FALSE}
train_df |>
  gg_tsdisplay(difference(TFR, 2), plot_type = 'partial')
```

```{r, echo = FALSE, message = FALSE}
train_df |>
  features(TLB_s, list(
    unitroot_kpss,
    unitroot_ndiffs)) |>
  mutate_if(is.numeric, ~ as.character(signif(.,3))) |>
  knitr::kable(caption = "Table 6: ")
```

```{r, echo = FALSE, message = FALSE}
train_df |>
  gg_tsdisplay(difference(TLB_s, 1), plot_type = 'partial')
```

#### Fitting

```{r, echo = FALSE, message = FALSE, warning = FALSE}
fit_TFR <-
  train_df |>
  model(TFR_021 = ARIMA(TFR ~ pdq(0,2,1)))

fit_TLB <-
  train_df |>
  model(
    TLB_010 = ARIMA(TLB_s ~ 1 + pdq(0,1,0)),
    TLB_110 = ARIMA(TLB_s ~ pdq(1,1,0))
  )
```

```{r, echo = FALSE, message = FALSE}
bind_rows(
  fit_TFR |>
  tidy(),
  fit_TLB |>
  tidy()) |>
  mutate_if(is.numeric, ~ as.character(signif(.,3))) |>
  knitr::kable(caption = "Table 7: ")
```

### Research Questions

\newpage

### Statistical Appendix

\newpage

### References

-   Singapore Department of Statistics. (2026). *Births and fertility rates, annual* \[Data set\]. data.gov.sg. <https://data.gov.sg/datasets/d_e39eeaeadb571c0d0725ef1eec48d166/view>
-   Singapore Department of Statistics. (2026). *Births, crude birth rate (CBR), total fertility rate (TFR) and reproduction rates* \[Data set\]. SingStat Table Builder. <https://tablebuilder.singstat.gov.sg/table/TS/M810091>
